---
title: "Assignment 1 - Language development in autistic and neurotypical children"
output: html_document
date: "2022-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(brms)

pacman::p_load(glue,
       data.table,
       moments,
       tidybayes,
       ggplot2,
       purr,
       ggridges,
       ellipse,
       cowplot,
       viridis)

```

```{r}
df <- read.csv("data_clean.csv") 
```


# Assignment 1  - Language development in autistic and neurotypical children

## Quick recap
Autism Spectrum Disorder is often related to language impairment. However, this phenomenon has rarely been empirically traced in detail: i) relying on actual naturalistic language production, ii) over extended periods of time.

We therefore videotaped circa 30 kids with ASD and circa 30 comparison kids (matched by linguistic performance at visit 1) for ca. 30 minutes of naturalistic interactions with a parent. We repeated the data collection 6 times per kid, with 4 months between each visit. We transcribed the data and counted: 
i) the amount of words that each kid uses in each video. Same for the parent.
ii) the amount of unique words that each kid uses in each video. Same for the parent.
iii) the amount of morphemes per utterance (Mean Length of Utterance) displayed by each child in each video. Same for the parent. 

This data is in the file you prepared in the previous class, but you can also find it here:https://www.dropbox.com/s/d6eerv6cl6eksf3/data_clean.csv?dl=0


## The structure of the assignment

We will be spending a few weeks with this assignment. In particular, we will:

Part 1) simulate data in order to better understand the model we need to build, and to better understand how much data we would have to collect to run a meaningful study (precision analysis)

Part 2) analyze our empirical data and interpret the inferential results

Part 3) use your model to predict the linguistic trajectory of new children and assess the performance of the model based on that.

As you work through these parts, you will have to produce a written document (separated from the code) answering the following questions:

Q1 - Briefly describe your simulation process, its goals, and what you have learned from the simulation. Add at least a plot showcasing the results of the simulation. Make a special note on sample size considerations: how much data do you think you will need? what else could you do to increase the precision of your estimates?

Q2 - Briefly describe the empirical data and how they compare to what you learned from the simulation (what can you learn from them?). Briefly describe your model(s) and model quality. Report the findings: how does development differ between autistic and neurotypical children (N.B. remember to report both population and individual level findings)? which additional factors should be included in the model? Add at least one plot showcasing your findings.

Q3 - Given the model(s) from Q2, how well do they predict the data? Discuss both in terms of absolute error in training vs testing; and in terms of characterizing the new kids' language development as typical or in need of support.


Below you can find more detailed instructions for each part of the assignment.

## Part 1 - Simulating data

Before we even think of analyzing the data, we should make sure we understand the problem, and we plan the analysis. To do so, we need to simulate data and analyze the simulated data (where we know the ground truth).

In particular, let's imagine we have n autistic and n neurotypical children. We are simulating their average utterance length (Mean Length of Utterance or MLU) in terms of words, starting at Visit 1 and all the way to Visit 6.
In other words, we need to define a few parameters:
- average MLU for ASD (population mean) at Visit 1 and average individual deviation from that (population standard deviation)
- average MLU for TD (population mean) at Visit 1 and average individual deviation from that (population standard deviation)
- average change in MLU by visit for ASD (population mean) and average individual deviation from that (population standard deviation)
- average change in MLU by visit for TD (population mean) and average individual deviation from that (population standard deviation)
- an error term. Errors could be due to measurement, sampling, all sorts of noise. 


Note that this makes a few assumptions: population means are exact values; change by visit is linear (the same between visit 1 and 2 as between visit 5 and 6). This is fine for the exercise. In real life research, you might want to vary the parameter values much more, relax those assumptions and assess how these things impact your inference.


We go through the literature and we settle for some values for these parameters:
- average MLU for ASD and TD: 1.5 (remember the populations are matched for linguistic ability at first visit)
- average individual variability in initial MLU for ASD 0.5; for TD 0.3 (remember ASD tends to be more heterogeneous)
- average change in MLU for ASD: 0.4; for TD 0.6 (ASD is supposed to develop less)
- average individual variability in change for ASD 0.4; for TD 0.2 (remember ASD tends to be more heterogeneous)
- error is identified as 0.2

This would mean that on average the difference between ASD and TD participants is 0 at visit 1, 0.2 at visit 2, 0.4 at visit 3, 0.6 at visit 4, 0.8 at visit 5 and 1 at visit 6.

With these values in mind, simulate data, plot the data (to check everything is alright); and set up an analysis pipeline.
Remember the usual bayesian workflow:
- define the formula
- define the prior
- prior predictive checks
- fit the model
- model quality checks: traceplots, divergences, rhat, effective samples
- model quality checks: posterior predictive checks, prior-posterior update checks
- model comparison

Once the pipeline is in place, loop through different sample sizes to assess how much data you would need to collect. N.B. for inspiration on how to set this up, check the tutorials by Kurz that are linked in the syllabus.

BONUS questions for Part 1: what if the difference between ASD and TD was 0? how big of a sample size would you need? What about different effect sizes, and different error terms?

1: simulate data
```{r}
#set the seed
set.seed(1)

#define number of participants for each group
n <- 50

# define the means at base level
mu_ASD <- 1.5
mu_TD <- 1.5

#define the standard deviations at base level
sd_ASD <- 0.5 
sd_TD <- 0.3

#define the means at change
mu_change_ASD <- 0.4
mu_change_TD <- 0.6

#define the standard deviations at change
sd_change_ASD <- 0.4
sd_change_TD <- 0.2

#define error
e <- 0.2

#define the simulations for intercept, slope and error (for both groups)
B0A <- rlnorm(n, mean = mu_ASD, sd = sd_ASD)
B0TD <- rlnorm(n, mean = mu_TD, sd = sd_TD)
B1A <- rlnorm(n, mean = mu_change_ASD, sd = sd_change_ASD)
B1TD <- rlnorm(n, mean = mu_change_TD, sd = sd_change_TD)
EA <- rlnorm(n, mean = 0, sd = e)
ETD <- rlnorm(n, mean = 0, sd = e)

#set up the data
sim_df <-
  tibble(group = rep(c("ASD", "TD"), each = n)) %>% #Add gruop n times
  mutate(intercept  = ifelse(group == "ASD", #Add intercept
                            B0A,
                            B0TD ))%>%
  mutate(slope = ifelse(group == "ASD", #Add slope
                            B1A ,
                            B1TD ))%>%
  mutate(error = ifelse(group == "ASD", #Add Error
                            EA,
                            ETD))%>%
  mutate(ID = row_number())%>% #Add ID
  slice(rep(1:n(), each = 6)) %>% #repeat each row 6 times
  add_column(visit=rep(c(1,2,3,4,5,6),times=n*2))  #add visit with numbers from 1 
 

for(i in seq(nrow(sim_df))) {
  sim_df$MLU[i] <- rnorm(1, sim_df$intercept[i]  +
                              sim_df$slope[i]  *
                            (sim_df$visit[i] -1))
  
}


#organize coloums in the right order
sim_df <- sim_df[, c(1, 5, 6, 2, 3, 4, 7)]


```


```{r}
#Visualize simulations
ggplot(sim_df, aes(visit, MLU, color = group, group = ID))+
  theme_bw()+
  geom_point()+
  geom_line(alpha
            =0.3)
```


```{r}
#Visualize simulations and calculate

d_intercept_v1_asd <- sim_df%>%
  filter(group == "ASD")%>%
  filter(visit== "1")%>%
  select(intercept)

hist(d_intercept_v1_asd$intercept)
print(mean(d_intercept_v1_asd$intercept))
print(sd(d_intercept_v1_asd$intercept))

d_intercept_v1_td <- sim_df%>%
  filter(group == "TD")%>%
  filter(visit== "1")%>%
  select(intercept)

hist(d_intercept_v1_td$intercept)
print(mean(d_intercept_v1_td$intercept))
print(sd(d_intercept_v1_td$intercept))

d_slope_change_asd <- sim_df%>%
  filter(group == "ASD")%>%
  filter(visit== "1")%>%
  select(slope)

hist(d_slope_change_asd$slope)
print(mean(d_slope_change_asd$slope))
print(sd(d_slope_change_asd$slope))

d_slope_change_td <- sim_df%>%
  filter(group == "TD")%>%
  filter(visit== "1")%>%
  select(slope)

hist(d_slope_change_td$slope)
print(mean(d_slope_change_td$slope))
print(sd(d_slope_change_td$slope))
```


```{r}
#Define the formula
MLU_f3 <- bf(MLU ~ 0 + group + group:visit + (1 + visit| ID))
```


```{r}
#Investigating what priors to set
get_prior(data = sim_df,
          family = gaussian,
          MLU_f3)
```

#Set prior

```{r}

MLU_p <- c(
  prior(normal(1.5, 0.4), class = b, coef = "groupASD"), #intercept foo ASD
  prior(normal(0, 0.5), class = b), #slope (for both), put to be the same for the group so we want the data to persuede us
  prior(normal(1.5, 0.4), class = b, coef = "groupTD"), #intercept for TD
  prior(normal(0, 0.5), class = sd), #prior for sd #calculated: mean(sim_df$MLU)- log(mean(sim_df$MLU)-sd(sim_df$MLU)) = 0.539 # take the mean of all MLU- minus what happens when you move 1sd away from this mean and take this at a log scale. 
  prior(lkj(1), class= cor),
  prior(normal(0, 0.5), class = sigma)
  )

#the sd for the 2 intercepts are set to the same (inspired by the littiature/previous studies) to again persuade us that there is a difference
#overall we 'assume' that the groups are the same and want the data to persuade us

```

```{r}
#Make model which samples only from the priors
#Be aware of family, maybe use log-normal

MLU_prior_m3 <- brm(
    MLU_f3,
    data = sim_df,
    family = gaussian,
    prior = MLU_p,
    sample_prior = "only", 
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20)
)
```
```{r}
#Prior predictive checks using pp-check
#pacman::p_load(extrafont)
#extrafont:::fonttable_file()
#bayesplot_theme_set(theme_default() + theme(text=element_text(family="")))
pp_check(MLU_prior_m3, ndraws = 100)
```




```{r}
#Plot the priors
#Produces a trace plot which shows the sampled parameter value at each iteration. If the distributions have converged, the plot should look like a hairy caterpillar

plot(MLU_prior_m3)

```


```{r}
#Investigating the parameters (Parameter recovery) of the model
#Interpret the results, look at CI
#Is there need for adjustments of priors

print(MLU_prior_m3)
```




```{r}
#Now: Fit a model that also samples from the simulation data and not only the prior

MLU_prior_m3_fit <- brm(
    MLU_f3,
    data = sim_df,
    family = gaussian,
    prior = MLU_p,
    sample_prior = T, 
    chains = 2,
    cores = 2,
    control = list(
      adapt_delta = 0.99,
      max_treedepth = 20)
)


```


```{r}
#pp-check on the new model that also samples both from the simulation data and the priors
#Check whether the plot has zoomed in compared to the first pp-check

#pacman::p_load(extrafont)
#extrafont:::fonttable_file()
#bayesplot_theme_set(theme_default() + theme(text=element_text(family="")))
pp_check(MLU_prior_m3_fit, ndraws = 100)
```

```{r}
variables(MLU_prior_m3_fit)
```


```{r}
#Prior posterior update checks on the new model
#Has the model learned?

#Sample the parameters of interest:
Posterior_m3 <- as_draws_df(MLU_prior_m3_fit)


#Plot the prior-posterior update plot for the intercept TD:
ggplot(Posterior_m3) +
  geom_density(aes(prior_b_groupTD), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_groupTD), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Intercept for TD') +
  theme_classic()


#Plot the prior-posterior update plot for the intercept ASD:
ggplot(Posterior_m3) +
  geom_density(aes(prior_b_groupASD), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_groupASD), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Intercept for ASD') +
  theme_classic()


#Plot the prior-posterior update plot for the Slope ASD:
ggplot(Posterior_m3) +
  geom_density(aes(`prior_b_groupASD:visit`), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(`b_groupASD:visit`), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Slope for ASD') +
  theme_classic()


#Plot the prior-posterior update plot for the Slope TD:
ggplot(Posterior_m3) +
  geom_density(aes(`prior_b_groupTD:visit`), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(`b_groupTD:visit`), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Slope for TD') +
  theme_classic()


#Plot the prior-posterior update plot for Sigma
ggplot(Posterior_m3) +
  geom_density(aes(prior_sigma), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sigma), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Sigma') +
  theme_classic()


#Plot the prior-posterior update plot for Correlation 
ggplot(Posterior_m3) +
  geom_density(aes(prior_cor_ID), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(cor_ID__Intercept__visit), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Correlation') +
  theme_classic()

```


```{r}
#Investigating the parameters (Parameter recovery) of the model
#Interpret the results, look at CI, do they include the prior?
#Is there need for adjustments of priors 

print(MLU_prior_m3_fit)
```



```{r}
#Try different sample sizes and see what effect they have
#Visualize this
#Our overall conclusion should be something like 50 n (according to Riccardo)
#Do include conceptual explanation/considerations: limitations due to e.g. funding, time

updated_fit <-
  update(MLU_prior_m3_fit,
         newdata = sim_df,
         seed = 2)

sim_function <- function(seed, n) {
  
  set.seed(seed)
  
 
  tibble(group = rep(c("ASD", "TD"), each = n)) %>% #Add gruop n times
  mutate(intercept  = ifelse(group == "ASD", #Add intercept
                            B0A,
                            B0TD ))%>%
  mutate(slope = ifelse(group == "ASD", #Add slope
                            B1A ,
                            B1TD ))%>%
  mutate(error = ifelse(group == "ASD", #Add Error
                            EA,
                            ETD))
}

sim_function(seed = 123, n = 2)


n_sim <- 100
t1 <- Sys.time() # this will help us track time


# here's the main event!
s <-
  tibble(seed = 1:n_sim) %>% 
  mutate(sim_df = map(seed, sim_function, n = 50)) %>% 
  mutate(MLU_prior_m3_fit = map2(sim_df, seed, ~update(MLU_prior_m3_fit, sim_df = .x, seed = .y)))

t2 <- Sys.time()

parameters <-
  s %>% 
  mutate(group = map(fit, ~ fixef(.) %>% 
                           data.frame() %>% 
                           rownames_to_column("parameter"))) %>% 
  unnest(group)


parameters %>% 
  select(-sim_df, -MLU_prior_m3_fit) %>% 
  filter(parameter == "group") %>% 
  head()


parameters %>% 
  filter(parameter == "group") %>% 
  
  ggplot(aes(x = seed, y = Estimate, ymin = Q2.5, ymax = Q97.5)) +
  geom_hline(yintercept = c(0, .5), color = "white") +
  geom_pointrange(fatten = 1/2) +
  labs(x = "seed (i.e., simulation index)",
       y = expression(beta[1]))

parameters %>% 
  filter(parameter == "group") %>% 
  mutate(check = ifelse(Q2.5 > 0, 1, 0)) %>% 
  summarise(power = mean(check))


```




# Part 2 - Strong in the Bayesian ken, you are now ready to analyse the actual data

- Describe your sample (n, age, gender, clinical and cognitive features of the two groups) and critically assess whether the groups (ASD and TD) are balanced. Briefly discuss whether the data is enough given the simulations in part 1.
- Describe linguistic development (in terms of MLU over time) in TD and ASD children (as a function of group). Discuss the difference (if any) between the two groups.
- Describe individual differences in linguistic development: do all kids follow the same path? Are all kids reflected by the general trend for their group?

- Include additional predictors in your model of language development (N.B. not other indexes of child language: types and tokens, that'd be cheating). Identify the best model, by conceptual reasoning, model comparison or a mix. Report the model you choose (and name its competitors, if any) and discuss why it's the best model.


```{r}

```


# Part 3 - From explanation to prediction

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=1
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=1
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=1

Relying on the model(s) you trained in part 2 of the exercise, create predictions for the test set and assess how well they do compared to the actual data.

- Discuss the differences in performance of your model in training and testing data. Is the model any good?
- Let's assume you are a speech therapy clinic. You want to assess whether the kids in your test sample will have a typical (like a TD) development, or they will have a worse one, in which case they should get speech therapy support. What do your predictions tell you about that? Which kids would you provide therapy for? Is the model any good?

```{r}


```
