---
title: "A3_Machine_learning"
author: "Ida Dencker"
date: "2022-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages

pacman::p_load(tidyverse, dplyr, tibble, parsnip, recipes, magrittr)

pacman::p_load(msm, tidyverse, brms, grid, gridExtra, readxl, metafor, tibble, ggplot2, hrbrthemes, reshape, dplyr, patchwork, hrbrthemes)
```

# The assignment

The Machine Learning assignment has 3 main parts: First we create a skeptical and an informed simulation, based on the meta-analysis. Second we build and test our machine learning pipeline on the simulated data. Third we apply the pipeline to the empirical data.

The report for the exam, thus, consists of the answer to all the following prompts:
- Describe your machine learning pipeline. Produce a diagram of it to guide the reader (e.g. see Rybner et al 2022 Vocal markers of autism: Assessing the generalizability of ML models), and describe the different parts: data budgeting, data preprocessing, model choice and training, assessment of performance.
- Briefly justify and describe your use of simulated data, and results from the pipeline on them.
- Describe results from applying the ML pipeline to the empirical data and what can we learn from them.

Remember: plots are very very important to communicate your process and results.

## Part I - Simulating data

Use the meta-analysis reported in Parola et al (2020), create a simulated dataset with 100 matched pairs of schizophrenia and controls, each participant producing 10 repeated measures (10 trials with their speech recorded). for each of these "recordings" (data points) produce 10 acoustic measures: 6 from the meta-analysis, 4 with just random noise. Do the same for a baseline dataset including only 10 noise variables. Tip: see the slides for the code. 


# packages 
```{r}
pacman::p_load(msm, tidyverse, brms, grid, gridExtra, readxl, metafor, dplyr, magrittr, reshape)
```
# (load data)
```{r}
data <- read_csv("Ass3_empiricalData1.csv")
```


`

-------
# Simulation
( først en chunk for df 1 (6 fra ma og 4 noise) og for df 2(baseline kun med noise)

#Simulating data (df 1 = with 6 from the meta-analysis, 4 with just random noise)

```{r}
#temp= temporarely
#wide= using the pivot wider function


# Define population size, we want 100 matched pairs, i.e 50 HC and 50 SZ
n <- 100
#Define trial
trials <- 10

# Define effect sizes
#see table 2 from the article for the numbers
InformedEffectMean <- c(0.25, -0.55, 0.74, -1.26, 0.05, 1.89, 0, 0, 0, 0)

# Define individual variability from population and across trails and measurement error
IndividualSD <-  1
TrialSD <- 0.5
Error <- 0.2

# Conditions
Schizophrenia <- rnorm(1, rnorm(1, 0.21, 0.5), 0.2)
Control <- rnorm(1, rnorm(1, -0.21, 0.5), 0.2)


#Creating the temporarily dataframe 
# For each pair of participant, we need to identity the true effect size
for(i in seq(10)) {
  temp_informed <- tibble(
    ID = seq(n),
    TrueEffect = rnorm(n, InformedEffectMean[i], IndividualSD),
    Variable = paste0("v", i))
  if(i == 1) {
    
    d_informed_true <- temp_informed
  } else {
    d_informed_true <- rbind(d_informed_true, temp_informed)
  }
}

# Create tibble with one row per trial
d_trial <- tibble(expand_grid(ID = seq(n), Trial = seq(trials), Group = c("Schizophrenia", "Control")))




d_informed <- merge(d_informed_true, d_trial)

for(i in seq(nrow(d_informed))){
  d_informed$measurement[i] <- ifelse(d_informed$Group[i]=="Schizophrenia",
                                      rnorm(1, rnorm(1, d_informed$TrueEffect[i]/2, TrialSD), Error),
                                      rnorm(1, rnorm(1, -d_informed$TrueEffect[i]/2, TrialSD), Error))
}

#Creating the final dataframe using the wide function on the d_informed dataset
d_informed_wide <- d_informed %>% 
  mutate(TrueEffect= NULL) %>% 
  pivot_wider(names_from = Variable,
              values_from = measurement)

```

#second simulation (df 2 = baseline dataset including only 10 noise variables)

# second simulation 
(df 2 = baseline dataset including only 10 noise variables)
```{r}
#... and now we do the same for the second data set (baseline dataset including only 10 noise variables)

skeptic_EffectMean <- rep(0,10)

for(i in seq(10)) {
  temp_skeptic <- tibble(
    ID = seq(n),
    TrueEffect = rnorm(n, skeptic_EffectMean[i], IndividualSD),
    Variable = paste0("v", i))
  if(i == 1) {
    d_informed_true <- temp_informed
    d_skeptic_true <- temp_skeptic
  } else {
    d_skeptic_true <- rbind(d_skeptic_true, temp_skeptic)
  }
}

d_skeptic <- merge(d_skeptic_true, d_trial)

for(i in seq(nrow(d_skeptic))){
  d_skeptic$measurement[i] <- ifelse(d_skeptic$Group[i]=="Schizophrenia",
                                      rnorm(1, rnorm(1, d_skeptic$TrueEffect[i]/2, TrialSD), Error),
                                      rnorm(1, rnorm(1, -d_skeptic$TrueEffect[i]/2, TrialSD), Error))
}

d_skeptic_wide <- d_skeptic %>%
  mutate(TrueEffect= NULL) %>% 
  pivot_wider(names_from = Variable,
              values_from = measurement)
```



## Part II - ML pipeline on simulated data

On the two simulated datasets (separately) build a machine learning pipeline: 

i) create a data budget (e.g. balanced training and test sets); 


```{r}
TestID <- sample(seq(n), 20)
                  


# Using the ! to say what is not in the first one put in the second one
train_informed <- d_informed_wide %>% subset(!(ID %in% TestID))
test_informed <- d_informed_wide %>% subset((ID %in% TestID))
```


ii) pre-process the data (e.g. scaling the features); 

#making ID and Trials as character - so they won't be scaled and centered in the next step
```{r}
d_informed_wide <- d_informed_wide %>% 
  mutate(ID = as.character(ID)) %>% 
  mutate(Trial = as.character(Trial))
```

```{r}
scaled_informed <- train_informed %>% 
  recipe(Group ~ .) %>% #define the outcome
  step_scale(all_numeric()) %>% #scales all the numeric values
  step_center(all_numeric()) %>% #centers all the numeric values
  prep(training = train_informed, retain = TRUE)

scaled_test_informed <- test_informed %>% 
  recipe(Group ~ .) %>% #define the outcome
  step_scale(all_numeric()) %>% #scales all the numeric values
  step_center(all_numeric()) %>% #centers all the numeric values
  prep(training = test_informed, retain = TRUE)
  
train_informed_scaled <- juice(scaled_informed)
test_informed_scaled <- juice(scaled_test_informed)

```


iii) fit and assess a classification algorithm on the training data (e.g. Bayesian multilevel logistic regression)

```{r}
#create formula
f1 <- bf(Group ~ 1 + v1 + v2 + v3 + v4 + v5 +v6 +v7 +v8 +v9 +v10)
```



```{r}
#get prios
get_prior(f1, train_informed_scaled, family=bernoulli)

```

Do include all 10 v, b.c. we want to know does he model apply to signal plus noise
#Use conservative prior
A lasso procedure: 


```{r}
#set priors
f1_prior <- c(
prior(normal(0, 1), class = Intercept),
prior(normal(0, 0.3), class = b) 
)
```

```{r}
# Build model

fitted_model_1 <- 
  brm(
    f1,
    data = train_informed_scaled,
    family = bernoulli,
    prior = f1_prior,
    sample_prior = T,
    iter = 2000, 
    warmup = 1000,
    cores = 2,
    chains = 2,
    refresh=0,
    control = list(
      adapt_delta = 0.999,
      max_treedepth = 20)
  )



```

```{r}
install.packages('RcppEigen')
```

```{r}
library(RcppEigen)

```

#look at pp-check of the fitted model
```{r}
#Posterior predictive checks

pp_check(fitted_model_1, ndraws = 100)
```


#Parameter recovery
```{r}
print(fitted_model_1)
```

#Prior posterior update checks 
```{r}

variables(fitted_model_1)
Posterior_f1 <- as_draws_df(fitted_model_1)

ggplot(Posterior_f1) +
  geom_density(aes(prior_Intercept), fill="chartreuse2", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="deeppink", color="black",alpha=0.6) + 
  xlab('Intercept') +
  theme_classic()+
  ggtitle("intercept for study effect size")+
      theme(plot.title = element_text(size = 10, face = "bold"))

#green = prior
#pink = simulated data 

```


iv) assess performance on the test set; 



v) discuss whether performance is as expected and feature importance is as expected.

Bonus question: replace the bayesian multilevel regression with a different algorithm, e.g. SVM or random forest (but really, anything you'd like to try).








## Part III - Applying the ML pipeline to empirical data

Download the empirical dataset from brightspace and apply your ML pipeline to the new data, adjusting where needed. Warning: in the simulated dataset we only had 10 features, now you have many more! Such is the life of the ML practitioner. Consider the impact a higher number of features will have on your ML inference, and decide whether you need to cut down the number of features before running the pipeline (or alternatively expand the pipeline to add feature selection).

Data: https://www.dropbox.com/s/7ky1axvea33lgye/Ass3_empiricalData1.csv?dl=0